---
title: "Pricing Trends in Airbnb Rentals in European Cities"
execute:
    echo: false
---

```{r,message=FALSE, warning=FALSE}
#libraries
library(tidyverse)
library(tidyr)
library(patchwork)
```

```{r}
#all data
combined_data <- read_csv("raw_csvs/airbnb_combined.csv",show_col_types = FALSE)

#data with no nas
cleaned_data <- combined_data |> drop_na()

avg_price_by_city <- cleaned_data |>
  group_by(city) |>
  summarize(realSum = mean(realSum, na.rm = TRUE)) |>
  ungroup()

#for sect. 1
avg_price_by_distance <- cleaned_data |>
  mutate(dist = cut(dist, 
    breaks = c(0, 5, 10, 15, 20, Inf),
    labels = c("0-5km", "5-10km", "10-15km", "15-20km", "20+km"),
    include.lowest = TRUE)) |>
  group_by(dist) |>
  summarize(realSum = mean(realSum, na.rm = TRUE)) |>
  ungroup()

#for sect 2.
avg_price_by_met_dist <- cleaned_data |>
  mutate(metro_dist = cut(metro_dist, 
    breaks = c(0, 3, 6, 9, 12, 15, Inf),
    labels = c("0-3km", "3-6km", "6-9km", "9-12km", "12-15km", "15+km"),
    include.lowest = TRUE)) |>
  group_by(metro_dist) |>
  summarize(realSum = mean(realSum, na.rm = TRUE)) |>
  ungroup()

#London and Amst only data
#If I had more time, to make my dataset more representative I would have adjusted this for the fact that there are many more Amsterdam data points than London data points
london_amsterdam <- cleaned_data |> filter(city %in% c("amsterdam", "london")) |> arrange(desc(city))

#Just Amst
amst <- london_amsterdam |> filter(city == c("amsterdam"))

#Just London
lon <- london_amsterdam |> filter(city == c("london"))

```

```{r}
#Main plot
avg_price_by_city|>ggplot(aes(x=city, y=realSum))+geom_col()+scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600))+labs(title="Average Price of Airbnb Rentals Across European Cities", subtitle="Rentals in Amsterdam appear to be significantly more expensive than renals in other countries.", y="Average Rental Price (in Euros)", x="City")
```

Over the last several years, Airbnb vacation rentals have been rising in popularity across all major European cities. The Airbnb listings in Amsterdam are much more expensive than rentals in other European cities, even  famously expensive cities like London. I sought to understand why rentals in Amsterdam are so expensive in comparision to rentals in other European cities- London in perticular. By examining trends in distance to city center, distance to the nearest metro station, number of bedrooms, and types of rentals, I determined this surprising differnce between the average price of rentals in  Amsterdam and London may be caused by London Airbnbs being smaller, farther from tourist destinations like the city center, and less convienently located by metro stations. However, it is important to note that one major reason for the difference in average rental price between these two cities could be that there are significantly more London Airbnb data points than Amsterdam data points, meaning the Lomdon data is more representative.
