---
title: "Airbnb Listings in Europe"
execute:
    echo: false
---

```{r,message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyr)
library(patchwork)
```

```{r}
combined_data <- read_csv("raw_csvs/airbnb_combined.csv",show_col_types = FALSE)

cleaned_data <- combined_data |> drop_na()

avg_price_by_city <- cleaned_data |>
  group_by(city) |>
  summarize(realSum = mean(realSum, na.rm = TRUE)) |>
  ungroup()

avg_price_by_distance <- cleaned_data |>
  mutate(dist = cut(dist, 
    breaks = c(0, 5, 10, 15, 20, Inf),
    labels = c("0-5km", "5-10km", "10-15km", "15-20km", "20+km"),
    include.lowest = TRUE)) |>
  group_by(dist) |>
  summarize(realSum = mean(realSum, na.rm = TRUE)) |>
  ungroup()

avg_price_by_met_dist <- cleaned_data |>
  mutate(metro_dist = cut(metro_dist, 
    breaks = c(0, 3, 6, 9, 12, 15, Inf),
    labels = c("0-3km", "3-6km", "6-9km", "9-12km", "12-15km", "15+km"),
    include.lowest = TRUE)) |>
  group_by(metro_dist) |>
  summarize(realSum = mean(realSum, na.rm = TRUE)) |>
  ungroup()

london_amsterdam <- cleaned_data |> filter(city %in% c("amsterdam", "london")) |> arrange(desc(city))

amst <- london_amsterdam |> filter(city == c("amsterdam"))

lon <- london_amsterdam |> filter(city == c("london"))

```

```{r}
avg_price_by_city|>ggplot(aes(x=city, y=realSum))+geom_col()+scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600))+labs(title="Average Price of Airbnb Rentals Across European Cities", subtitle="Rentals in Amsterdam appear to be significantly more expensive than renals in other countries.", y="Average Rental Price (in Euros)", x="City")
```

```{r,message=FALSE}
avg_price_by_distance|>ggplot(aes(x=dist,y=realSum,group=1))+geom_col()+scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250, 300))+labs(x="Distance from City Center (in Kiolmeters)", y="Average Rental Price (in Euros)", title="Average Price of Airbnb Rentals in European Cities According to Distance from City Center", subtitle="Generally rentals closer to the city center are more expensive.")

cleaned_data|>ggplot(aes(x=dist,y=realSum,group=1))+geom_jitter(aes(color=city))+scale_y_log10()+geom_smooth(se = FALSE)+labs(title="Airbnb Rental Price in European Cities by Distance from City Center", subtitle= "More expensive rentals tend to be located closer to the city center than farther from it.",y="Rental Price (in Euros)", x="Distance from City Center (in Kiolmeters)")

plot1<- london_amsterdam|>ggplot(aes(x=dist,y=realSum, color=city))+geom_point(alpha=0.3)+scale_y_log10()+geom_smooth(method = "loess", formula = y ~ x, se = FALSE) + geom_smooth(method = "loess", formula = y ~ x, se = FALSE, linewidth = 1.5)+labs(title="Airbnb Rental Price in Amsterdam and London by Distance from City Center", subtitle="Rentals in Amsterdam and generally closer to the city center and more expensive", y="Rental Price (in Euros)", x="Distance from City Center (in Kiolmeters)")

plot2<- amst|>ggplot(aes(x=dist,y=realSum))+geom_point(alpha=0.3, color = "#F8766D")+scale_y_log10()+geom_smooth(method = "loess", formula = y ~ x, se = FALSE)+labs(y="Rental Price (in Euros)", x="Distance from City Center (in Kiolmeters)",title="Price of Airbnb Rentals in Amsterdam Relative to their Distance from the City Center", subtitle= "More expensive rentals tend to be located closer to the center of Amsterdam.")

plot3<-lon|>ggplot(aes(x=dist,y=realSum))+geom_point(alpha=0.3, color = "#00BFC4")+scale_y_log10()+geom_smooth(method = "loess", formula = y ~ x, se = FALSE)+labs(y="Rental Price (in Euros)", x="Distance from City Center (in Kiolmeters)", title="Price of London Airbnbs According to their Distance from the City Center", subtitle= "More expensive rentals tend to be located closer to the city center.")

plot1 
plot2 
plot3

```

```{r,message=FALSE}
avg_price_by_met_dist|>ggplot(aes(x=metro_dist,y=realSum,group=1))+geom_col()+scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250, 300))+labs(x="Distance from Nearest Metro Station Center (in Kiolmeters)", y="Average Rental Price (in Euros)", title="Average Price of Airbnb Rentals in European Cities According to Distance from the Nearest Metro Station", subtitle="More expensive Airbnbs tend to be closer to metro stations.")

cleaned_data|>ggplot(aes(x=metro_dist,y=realSum,group=1))+geom_jitter(aes(color=city))+scale_y_log10()+geom_smooth(se = FALSE)+labs(title="Airbnb Rental Price in European Cities by Distance from Nearest Metro Station", subtitle= "Rentals near metro stations tend to be more expensive than those farther from stations.",y="Rental Price (in Euros)", x="Distance from Nearest Metro Station (in Kiolmeters)")

plot4<- london_amsterdam|>ggplot(aes(x=metro_dist,y=realSum, color=city))+geom_point(alpha=0.3)+scale_y_log10()+geom_smooth(method = "loess", formula = y ~ x, se = FALSE) + geom_smooth(method = "loess", formula = y ~ x, se = FALSE, linewidth = 1.5)+labs(y="Rental Price (in Euros)", x="Distance from Nearest Metro Station (in Kiolmeters)",title="Price Airbnbs in Amsterdam and London Relative to their Distance from the Nearest Metro Station", subtitle= "Rentals in Amsterdam tend to be closer to metro stations and more expensive than those in London.")

plot5<- amst|>ggplot(aes(x=metro_dist,y=realSum))+geom_point(alpha=0.3, color = "#F8766D")+scale_y_log10()+geom_smooth(method = "loess", formula = y ~ x, se = FALSE)+labs(y="Rental Price (in Euros)", x="Distance from the Nearest Metro Station (in Kiolmeters)", title="Price of Amsterdam Airbnbs Relative to their Distance from the Nearest Metro Station", subtitle= "More expensive rentals tend to be located closer to metro stations.")

plot6<-lon|>ggplot(aes(x=metro_dist,y=realSum))+geom_point(alpha=0.3, color = "#00BFC4")+scale_y_log10()+geom_smooth(method = "loess", formula = y ~ x, se = FALSE)+labs(y="Rental Price (in Euros)", x="Distance from the Nearest Metro Station (in Kiolmeters)", title="Price of Airbnbs in London According to their Distance from the Nearest Metro Station", subtitle= "Rentals closer to metro stations tend to be more expensive.")

plot4
plot5
plot6
```

```{r}
cleaned_data %>%
  filter(!is.na(bedrooms), !is.na(realSum), realSum < 1000) %>%
  ggplot(aes(x = factor(bedrooms), y = realSum)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(0, 1600, by = 200)) +
  labs(
    x = "Number of Bedrooms",
    y = "Rental Price (in Euros)",
    title = "Distribution of Airbnb Rental Prices in European Cities According Number of Bedrooms",
    subtitle= "Ignoring outliers, more expensive rentals have more bedrooms."
  ) +
  theme_minimal()

london_amsterdam %>%
  filter(!is.na(bedrooms), bedrooms <= 6) %>%
  ggplot(aes(x = bedrooms, fill = city)) +
  geom_histogram(binwidth = 1, position = "dodge", color = "white") +
  labs(
    title = "Distribution of Airbnb Prices According Number of Bedrooms in Rentals in London and Amsterdam",
    x = "Number of Bedrooms",
    y = "Number of Rentals",
    subtitle= "Compared to rentals in Amsterdam, a disproportionate amount of rentals in London have fewer bedrooms, implying London rentals may be generally cheaper than Amsterdam Rentals because they have fewer bedrooms."
  ) +
  theme_minimal()

amst|> ggplot(aes(y=bedrooms)) + geom_bar() +labs(y="Number of Bedrooms", x="Number of Rentals",title="Distribution of Airbnb Prices in Amsterdam According Number of Bedrooms", subtitle="Most rentals in Amsterdam have one bedroom.") 
lon|> ggplot(aes(y=bedrooms)) + geom_bar()+labs(y="Number of Bedrooms", x="Number of Rentals",title="Distribution of Airbnb Prices in London by the Number of Bedrooms in the Rentals", subtitle="Most rentals in London have one bedroom.")
```

```{r}
cleaned_data %>%
  filter(!is.na(room_type), !is.na(realSum), realSum < 1000) %>%
  ggplot(aes(x = factor(room_type), y = realSum)) +
  geom_boxplot() +
  scale_y_continuous(breaks = seq(0, 1600, by = 200)) +
  theme_minimal() +
  labs(
    x = "Rental Type",
    y = "Rental Price (in Euros)",
    title = "Distribution of Airbnb  Prices in European Cities According Type of Rental",
    subtitle= "Entire home/apt rentals have higher medians than other rental types.")
london_amsterdam %>%
  ggplot(aes(x = factor(room_type), fill = city)) +
  geom_bar(position = "dodge") +
  theme_minimal() +
  labs(
    title = "Distribution of Airbnb Prices by Rental Type for Listings in London and Amsterdam",
    x = "Rental Type",
    y = "Number of Rentals",
    subtitle= "A smaller proportion of rentals in London are entire homes/apts, which tend to be more expensive than other rental types, than in Amsterdam, in part explaining why rentals in London tend to be more expensive than those in Amsterdam."
  )
amst|> ggplot(aes(y=room_type)) + geom_bar() + labs(y="Rental Type", x="Number of Rentals",title="Distribution of Airbnb Prices in Amsterdam According to Rental Type", subtitle="Most rentals in Amsterdam are entire homes/apts.")
lon|> ggplot(aes(y=room_type)) + geom_bar() + labs(y="Rental Type", x="Number of Rentals",title="Distribution of Airbnb Prices in London by Rental Type", subtitle="Most rentals in London are private rooms.")
```

Over the last several years Airbnb vacation rentals have been rising in popularity across all major European cities. Using Airbnb listing data compiled by researchers at the University of Warsaw, I seek to understand the factors contributing high guest satisfaction ratings and how those factors differ across European cities. The Airbnb listings in Amsterdam are much more expensive than rentals in other European cities; even in nutouriously expensive cities like London. By examining trends in distance to city center, distance to the nearest metro station, number of bedrooms, and types of rentals, I determined this surprising differnce between the average price of rentals in  Amsterdam and London can be attributed to rentals in London being smaller, farther from tourist destinations like the city center, and less convienently located by metro stations. 
